Supporting Information

### 2018-03-16 SupportingInformation.Rmd
adding Ext_risk_function

source("mtime.R") #A function needed for caching
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("normal_cop_AICw.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
source("TestStats_multivar_dataset.R")
source("skewness_multivar_dataset.R")
```

@@

Here, we want to test our $3$ types of nonparametric tail dependent statistics on both small $(35)$ and large $(1000)$ number of datapoints drawn from $4$ types of bivariate copula, namely, *Clayton* , *Normal*, *Frank* and *Survival Clayton*. Whereas *Normal* and *Frank* have symmetric tail dependence, *Clayton* and *Survival Clayton* have lower and upper tail dependence, respectively.
First we chose a set of parameters [using *iRho* from `Copula` package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$]. Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated $(Corl , Coru, Pl, Pu, D2l, D2u)$ stats. We took mean of our statistic $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation. We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [Figure \ref{fig_stat_testing1000}]. Statistic $Corl - Coru$, $Pl-Pu$ and $D2u-D2l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*. To verify this, we executed t-test and report the $p$-values for each case and most of the case shows $p>0.05$.

@@

```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"))}
set.seed(seed)
resultsloc<-"./Results/stat_results/stat_testing/"
b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
```

@@

#kendall, D2
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
mtext(side=1,line=1,text="Diagonal slice")
mtext(side=2,line=1,text=expression(D[2]))
axis(side=1,labels=F)
lines(x,D2stats_K[1,],type='l',lty='dotted')
lines(x,D2stats_K[2,],type='l',lty='dashed')
lines(x,D2stats_K[3,],type='l',lty='dashed')
lines(x,D2stats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_K,ylimits_D2)
vlineplot(x,ylimits_D2)

@@

<!--Now make and display a table, this one about asymmetry of tail dependence-->
```{r plot_soilCN_asymmtry_taildep, echo=F}
atdres<-data.frame(Statistic=c("Corl-Coru","Pl-Pu","D2u-D2l"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)
fractxt<-function(fracval)
{
  ptxt<-''
  if (fracval>.5*numsurrog)
  {
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog)
  {
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}
atdres[1,2]<-fractxt(corlmcoru_frac_K)
atdres[1,3]<-fractxt(corlmcoru_frac_S)
atdres[2,2]<-fractxt(PlmPu_frac_K)
atdres[2,3]<-fractxt(PlmPu_frac_S)
atdres[3,2]<-fractxt(D2umD2l_frac_K)
atdres[3,3]<-fractxt(D2umD2l_frac_S)
```

@@

```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count)}
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

@@

```{r tab_count_resamp,echo=F,results="markup"}
knitr::kable(stat_aphid_count$numericdf,
             format = "pandoc",
             caption = "Aphid_count : resampling result for stat indicating lower tail dependence\\label{tab_count_resamp}",
             booktabs=TRUE)
```

@@

![Corl-Coru vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![Pl-Pu vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![D2u-D2l vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_count_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero. $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.

@@

```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff)}
set.seed(seed)
resloc<-"./Results/stat_results/stat_aphid_ff/"

@@

```{r tab_ff_resamp,echo=F,results="markup"}
knitr::kable(stat_aphid_ff$numericdf,
             format = "pandoc",
             caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence\\label{tab_ff_resamp}",
             booktabs=TRUE)
```

@@

![Corl-Coru vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![Pl-Pu vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![D2u-D2l vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_ff_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_ff_resamp} indicates the fraction of resampled mean greater then zero. $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data. This proves that there is a significant upper tail dependence in *Leaf-curling plum aphid*-firstflight data.

@@

### Non-paramtric analysis of tail dependence  \label{Ex_pns_NPA}
```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea)}
set.seed(seed)
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

@@

```{r tab_pns_resamp,echo=F,results="markup"}
knitr::kable(stat_plankton_north_sea$numericdf,
             format = "pandoc",
             caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence\\label{tab_pns_resamp}",
             booktabs=TRUE)
```

@@

![Corl-Coru vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Corl-Coru}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![Pl-Pu vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![D2u-D2l vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_pns_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_pns_resamp} indicates the fraction of resampled mean greater then zero. $Corl-Coru$, $Pl-Pu$ and $D2u-D2l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.

@@

![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=50%, height=50% }

@@

![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=50%, height=50% }

@@

![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=50%, height=50% }

## Extinction risk

@@

>>>>

source("mtime.R") #A function needed for caching
source("getcopula.R")
source("OurBiCopSelect.R")
source("FittingCopula_selective_loc.R")
source("normal_cop_AICw.R")
source("aphid_data_calling.R")
source("plankton_north_sea_data_calling.R")
source("NonParamStat.R")
source("copsurrog2d.R")
source("TestStats_multivar_dataset.R")
source("skewness_multivar_dataset.R")
source("ExtremeTailDep.R")
```

@@

Here, we want to test our $3$ types of nonparametric tail dependent statistics on both small $(35)$ and large $(1000)$ number of datapoints drawn from $4$ types of bivariate copula, namely, *Clayton* , *Normal*, *Frank* and *Survival Clayton*. Whereas *Normal* and *Frank* have symmetric tail dependence, *Clayton* and *Survival Clayton* have lower and upper tail dependence, respectively.
First we chose a set of parameters [using *iRho* from `Copula` package] for each type of copula for which spearman correlation will be $10$ equispaced number in between [$0,0.9$]. Then for each parameter (or spearman value) of each type of copula we carried out $100$ simulation and calculated ($\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, $\Dsq_l$, $\Dsq_u$) stats. We took mean of our statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ over $100$ simulations and In Figure \ref{fig_stat_testing35}, we plotted those statistic against spearman correlation. We put the standard error bar around each point in the plot and these errors got minimized with increasing numbe of data points [Figure \ref{fig_stat_testing1000}]. Statistic $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ should be $0$ for spearman=$0$ for *Clayton* and *Survival Clayton* and should remain $0$ for all values of spearman for symmetric copula like *Normal* and *Frank*. To verify this, we executed t-test and report the $p$-values for each case and most of the case shows $p>0.05$.

@@

```{r stat_testing_plot, results='hide', echo=F, cache=T, cache.extra=list(seed,callfn,mtime("StatTestNumericResults.RData"),mtime("TestStats_multivar_dataset.R"))}
set.seed(seed)
resultsloc<-"./Results/stat_results/stat_testing/"
b<-callfn$paramlist
plotter_stat_testing(res_pt35_rankF,"res_pt35_rankF",b)
plotter_stat_testing(res_pt35_rankT,"res_pt35_rankT",b)
plotter_stat_testing(res_pt1000_rankF,"res_pt1000_rankF",b)
plotter_stat_testing(res_pt1000_rankT,"res_pt1000_rankT",b)
```

@@

#kendall, D2
par(fig=c(ywd/totwd,
          (ywd+panwd)/totwd,
          (xht)/totht,
          (xht+panht)/totht),
    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25,new=T)
plot(x,D2stats_d,type='l',col="red",xlim=xlimits,ylim=ylimits_D2)
mtext(side=1,line=1,text="Diagonal slice")
mtext(side=2,line=1,text=expression(D^{2}))
axis(side=1,labels=F)
lines(x,D2stats_K[1,],type='l',lty='dotted')
lines(x,D2stats_K[2,],type='l',lty='dashed')
lines(x,D2stats_K[3,],type='l',lty='dashed')
lines(x,D2stats_K[4,],type='l',lty='dotted')
text(xlimits[1],ylimits_D2[1],labels='E',cex=1.5,adj=c(.5,0))
fracplot(D2stats_frac_K,ylimits_D2)
vlineplot(x,ylimits_D2)

@@

<!--Now make and display a table, this one about asymmetry of tail dependence-->
```{r plot_soilCN_asymmtry_taildep, echo=F}
atdres<-data.frame(Statistic=c("$cor_l-cor_u$","$P_l-P_u$","$D^2_u-D^2_l$"),
                   Kendall=rep("",3),
                   Spearman=rep("",3),stringsAsFactors = F)
fractxt<-function(fracval)
{
  ptxt<-''
  if (fracval>.5*numsurrog)
  {
    ptxt<-paste(">",fracval,sep='')
  } 
  if (fracval<=.5*numsurrog)
  {
    ptxt<-paste("<",numsurrog-fracval,sep='')
  }
  return(ptxt)
}
atdres[1,2]<-fractxt(corlmcoru_frac_K)
atdres[1,3]<-fractxt(corlmcoru_frac_S)
atdres[2,2]<-fractxt(PlmPu_frac_K)
atdres[2,3]<-fractxt(PlmPu_frac_S)
atdres[3,2]<-fractxt(D2umD2l_frac_K)
atdres[3,3]<-fractxt(D2umD2l_frac_S)
```

@@

```{r stat_aphid_count, echo=F, cache=T, cache.extra=list(seed,data_aphid_count,mtime("NonParamStat.R"))}
set.seed(seed) 
resloc<-"./Results/stat_results/stat_aphid_count/"

@@

```{r tab_count_resamp,echo=F,results="markup"}
s<-stat_aphid_count$numericdf[,2:5]
rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
knitr::kable(s,
             format = "pandoc",
             caption = "Aphid_count : resampling result for stat indicating lower tail dependence\\label{tab_count_resamp}",
             booktabs=TRUE)
```

@@

![$cor_l$-$cor_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Corl-Coru}](./Results/stat_results/stat_aphid_count/Sp_10_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![$P_l$-$P_u$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_Pl-Pu}](./Results/stat_results/stat_aphid_count/Sp_10_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-count showing lower tail dependence\label{fig_count_D2u-D2l}](./Results/stat_results/stat_aphid_count/Sp_10_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_count_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_count_resamp} indicates the fraction of resampled mean greater then zero. $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that *Green spruce aphid*-count data has significant lower tail dependence.

@@

```{r stat_aphid_ff, echo=F, cache=T, cache.extra=list(seed,data_aphid_ff,mtime("NonParamStat.R"))}
set.seed(seed)
resloc<-"./Results/stat_results/stat_aphid_ff/"

@@

```{r tab_ff_resamp,echo=F,results="markup"}
s<-stat_aphid_ff$numericdf[,2:5]
rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
knitr::kable(s,
             format = "pandoc",
             caption = "Aphid_firstflight : resampling result for stat indicating upper tail dependence\\label{tab_ff_resamp}",
             booktabs=TRUE)
```

@@

![$cor_l$-$cor_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Corl-Coru}](./Results/stat_results/stat_aphid_ff/Sp_11_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![$P_l$-$P_u$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_Pl-Pu}](./Results/stat_results/stat_aphid_ff/Sp_11_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![$D^2_u$-$D^2_l$ vs. Distance plot for aphid-firstflight showing upper tail dependence\label{fig_ff_D2u-D2l}](./Results/stat_results/stat_aphid_ff/Sp_11_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_ff_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_ff_resamp} indicates the fraction of resampled mean greater then zero. $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with negative mean has all resampled mean less than $0$ for *Leaf-curling plum aphid*-firstflight data. This proves that there is a significant upper tail dependence in *Leaf-curling plum aphid*-firstflight data.

@@

### Non-paramtric analysis of tail dependence  \label{Ex_pns_NPA}
```{r stat_plankton_north_sea, echo=F, cache=T, cache.extra=list(seed,data_plankton_north_sea,mtime("NonParamStat.R"))}
set.seed(seed)
resloc<-"./Results/stat_results/stat_plankton_north_sea/"

@@

```{r tab_pns_resamp,echo=F,results="markup"}
s<-stat_plankton_north_sea$numericdf[,2:5]
rownames(s)<-c("spear","kend","cor~l~","cor~u~","P~l~","P~u~","$D^2$~l~","$D^2$~u~","cor~l~ - cor~u~","P~l~ - P~u~","$D^2$~u~ - $D^2$~l~")
knitr::kable(s,
             format = "pandoc",
             caption = "Plankton_north_sea : resampling result for stat indicating lower tail dependence\\label{tab_pns_resamp}",
             booktabs=TRUE)
```

@@

![$cor_l$-$cor_u$ vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Corl-Coru}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Corl-Coru_vs_D.pdf){ width=50%, height=50% }

![$P_l$-$P_u$ vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_Pl-Pu}](./Results/stat_results/stat_plankton_north_sea/Sp_16_Pl-Pu_vs_D.pdf){ width=50%, height=50% }

![$D^2_u$-$D^2_l$ vs. Distance plot for Plankton_north_sea showing lower tail dependence\label{fig_pns_D2u-D2l}](./Results/stat_results/stat_plankton_north_sea/Sp_16_D2u-D2l_vs_D.pdf){ width=50%, height=50% }

Table \ref{tab_pns_resamp} shows the mean of each non-parametric statistic taken from all pairwise locations and then based on a resampling results how the resampled mean distributed from or around the real mean. Fourth column named $fracgt0$ in Table \ref{tab_pns_resamp} indicates the fraction of resampled mean greater then zero. $\cor_l-\cor_u$, $\Ps_l-\Ps_u$ and $\Dsq_u-\Dsq_l$ statistic with positive mean and almost $100$% $fracgt0$ proves that plankton species data for *Ceratium furca* has significant lower tail dependence.

@@

![Skewness_count for species : Green spruce aphid\label{fig_skewness_count}](./Results/skewness_results/skewness_aphid_count/skewness_aphid_count_sp_10.pdf){ width=40%, height=40% }

@@

![Skewness_firstflight for species : Leaf-curling plum aphid\label{fig_skewness_ff}](./Results/skewness_results/skewness_aphid_ff/skewness_aphid_ff_sp_11.pdf){ width=40%, height=40% }

@@

![Skewness_plankton_north_sea for species : Ceratium_furca\label{fig_skewness_pns}](./Results/skewness_results/skewness_plankton_north_sea/skewness_plankton_north_sea_sp_16.pdf){ width=40%, height=40% }

## Extinction risk
```{r fn_popsim_ml_D, echo=F}

# A function to generate dispersal matrix D (a square matrix : numlocs by numlocs)
Dispersal_mat<-function(numlocs,d){
  D<-matrix(0,numlocs,numlocs)
  # A companion matrix (delta) that indicates how "off" a diagonal is:
  mat <- matrix(seq_len(numlocs ^ 2), ncol = numlocs)
  delta <- row(mat) - col(mat)
  #print(delta)

  diag(D)<-(1-d)
  D[delta ==1 | delta==-1] <- d/2

  D[numlocs-1,numlocs]<-d
  D[2,1]<-d

  stopifnot(all(colSums(D)==1)) # this is a check
  return(D)
}
#-----------------------------------------------------------------------------------
#The simulator of the multi-location model with a dispersal matrix.
#-------Args--------------------------------------------------
#p0       A length numlocs vector holding initial populations
#ns       An numsims by numlocs by numsteps array of the epsilons, 
#           where numsteps is the number of time steps you want
#
#--------Output------------------------------------------------------
#A numsims by numlocs by numsteps+1 array of populations
#----------------------------------------------------------------------
popsim_ml_D<-function(p0,ns,D,r,K){
  numsims<-dim(ns)[1]
  numlocs<-dim(ns)[2]
  numsteps<-dim(ns)[3]

  #initialization
  res<-array(NA,c(numsims,numlocs,numsteps+1))
  res[,,1]<-rep(p0,each=numsims)

  #step the populations forward
  for (tct in 1:numsteps){

    #stochastic ricker model : if r=0 then back to LC model

    #growth rates based on the noise
    lam_sto<-exp(ns[,,tct])  #numsims by numlocs matrix
    lam_ricker<-exp(r*(1-(res[,,tct]/K)))

    #growth prior to dispersal
    res[,,tct+1]<-lam_ricker*lam_sto*res[,,tct] #numsims by numlocs matrix

    #after dispersal
    tres<-t(res[,,tct+1]) # numlocs by numsims matrix after taking transpose
    temp<-D%*%tres  # a matrix multiplication (numlocs by numlocs) * (numlocs by numsims)
    res[,,tct+1]<-t(temp)

    #see if there is extinction
    h<-res[,,tct+1]
    h[h<1]<-0
    res[,,tct+1]<-h
  }

# function to calculate extinction risk
extrisk<-function(sims){
  totpop<-apply(FUN=sum,X=sims,MARGIN=c(1,3))
  return(apply(FUN=sum,X=(totpop==0),MARGIN=2)/dim(sims)[1])
}
```

```{r fn_plotter_ext_risk, echo=F}
# function to optionally plot extinction risk agsinst time and give you back extinction risk after specified numsteps 
plotter_ext_risk<-function(numsims,numsteps,numlocs,d,r,K,ploton){
  D<-Dispersal_mat(numlocs = numlocs,d=d)
  ns1<-retd(n=numsteps*numsims,d=numlocs,rl=1)# a righttail dep matrix(numpoints by numlocs,     
#                                                      numpoints=numsteps*numsims)

  ns1<-array(ns1,c(numsteps,numsims,numlocs))# convert to an array (numsteps by numsims by numlocs)
  ns1<-aperm(ns1,c(2,3,1)) # convert to an array (numsims by numlocs by numsteps)
  pops1<-popsim_ml_D(p0=rep(25,numlocs),ns=ns1,D=D,r=r,K=K)
  risk_right<-extrisk(pops1)

  ns2<-(-ns1)
  pops2<-popsim_ml_D(p0=rep(25,numlocs),ns=ns2,D=D,r=r,K=K)
  risk_left<-extrisk(pops2)

  if(ploton==T){
   par(mfrow=c(2,1))
    plot(0:(length(risk_right)-1),risk_right,type='l',
         xlab='Time',ylab='Risk',col='blue',panel.first = grid())
    mtext("blue : right tail-dep")
    plot(0:(length(risk_left)-1),risk_left,type='l',xlab='Time',ylab='Risk',col='red',panel.first = grid())
    mtext("red : left tail-dep")
  }

  return(list(risk_right_after_numsteps=risk_right[numsteps+1],
              risk_left_after_numsteps=risk_left[numsteps+1]))

}

```{r fn_varying_d, echo=F}
varying_d<-function(numsims,numsteps,numlocs,r,K){
  risk_right<-c()
  risk_left<-c()
  d_seq<-seq(from=0,to=1,by=0.01)
  for(d in d_seq){
    riskrl<- plotter_ext_risk(numsims=numsims,numsteps = numsteps,numlocs = numlocs,d=d,r=r,K=K,ploton=F)
    risk_r<-riskrl$risk_right_after_numsteps
    risk_l<-riskrl$risk_left_after_numsteps
    risk_right<-c(risk_right,risk_r)
    risk_left<-c(risk_left,risk_l)
  }
  par(mfrow=c(1,2))
  plot(d_seq,risk_right,xlab='d',ylab='Risk_right',ylim=c(0,max(risk_right)),type="b",col="blue",panel.first = grid())
  plot(d_seq,risk_left,xlab='d',ylab='Risk_left',ylim=c(0,max(risk_left)),type="b",col="red",panel.first = grid())
  #mtext("red : left tail-dep, blue : right tail-dep")
}
```

```{r fig_risk_K50, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=50, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K50}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=50)
```

```{r fig_risk_K100, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K100}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
```

```{r fig_risk_K200, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=200, numlocs=10, numsteps=25, numsims=1000.\\label{fig_risk_K200}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=200)
```

```{r fig_risk_numlocs5, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=5, numsteps=25, numsims=1000.\\label{fig_risk_numlocs5}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=5,r=1,K=100)
```

```{r fig_risk_numlocs15, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=15, numsteps=25, numsims=1000.\\label{fig_risk_numlocs15}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=10,r=1,K=100)
```

```{r fig_risk_numlocs25, echo=F, fig.cap="Plot Extinction risk vs. d(degree of dispersal) for r=1, K=100, numlocs=25, numsteps=25, numsims=1000.\\label{fig_risk_numlocs25}", fig.height=4, fig.width=8}
varying_d(numsims=1000,numsteps =25,numlocs=15,r=1,K=100)
```

  * With increasing $K$, extinction risk decreases [for a fixed $d$ , $r$ and $numlocs$] but this effect is more prominent in
case of left tail-dep data. [see Figure \ref{fig_risk_K50} - Figure \ref{fig_risk_K200}]

  * With increasing $numlocs$, extinction risk decreases [for a fixed $d$ , $r$ and $K$] but this effect is more prominent in
case of left tail-dep data. [see Figure \ref{fig_risk_numlocs5} - Figure \ref{fig_risk_numlocs25}]

\newpage

@@

<!--do a TL analysis on the data just created-->
```{r tl_analysis, echo=F, cache=T, cache.extra=list(mtime("TaylorsLawTools.R"),seed,mtime("copsurrognd.R"))}
#source("TaylorsLawTools.R")

#We are only doing spatial taylor's law because we know a priori temporal
#won't give anything interesting for the surrogates we use

#kend.stats.res<-list()
#kend.points.res<-list()
#spea.stats.res<-list()
#spea.points.res<-list()
#for (counter in 1:length(kend.surrog))
#{
 # ks<-kend.surrog[[counter]]
#  ss<-spea.surrog[[counter]]
 # kend.stats.res[[counter]]<-matrix(NA,5,numdat)
#  spea.stats.res[[counter]]<-matrix(NA,5,numdat)
 # kend.points.res[[counter]]<-array(NA,c(N,2,numdat))
  #spea.points.res[[counter]]<-array(NA,c(N,2,numdat))
  #for (sc in 1:numdat)
  #{
   # khold<-tl_stats(ks[,,sc])
  #  kend.stats.res[[counter]][,sc]<-khold$s.stats #spatial TL only
   # rownames(kend.stats.res[[counter]])<-names(khold$s.stats)
    #kend.points.res[[counter]][,,sc]<-as.matrix(khold$s.points)

    #shold<-tl_stats(ss[,,sc])
    #spea.stats.res[[counter]][,sc]<-shold$s.stats #spatial TL only
    #rownames(spea.stats.res[[counter]])<-names(shold$s.stats)
    #spea.points.res[[counter]][,,sc]<-as.matrix(shold$s.points)
  #}
#}
#names(kend.stats.res)<-names(kend.surrog)
#names(kend.points.res)<-names(kend.surrog)
#names(spea.stats.res)<-names(spea.surrog)
#names(spea.points.res)<-names(spea.surrog)

#DAN: now you want to compare whether tl held, and the slope (statistically and visually)
#across the different elements of kend.XX.res and spea.XX.res. See drawing of a figure layout on desk.
